Multi-line Intensity Mapping

### 2019-07-15 paper/ms.tex
Add disclaimer about shot noise
Update to reflect angle-averaged case

The intrinsic emission from a line $i$ is given by
\beq\label{eq:int_ps}
P_{i,i}(k, \mu, z) = \avg{I_i}^2 \avg{b_i}^2 (1 + \beta_i \mu^2)^2 
              D\left[\mu k \sigma_p(z)\right] P_{\delta, \delta}(k, z)\text{,}
              % + P_{i, \text{shot}}\text{,}
\eeq
where $\mu=k_{\parallel}/k$ is the cosine of the angle between the wavevector
$\bm{k}$ and the line of sight direction, $\avg{I_i}$ is the average specific
intensity of line $i$, and $\avg{b_i}$ is the average luminosity-weighted bias
of the emitting galaxies. The Kaiser effect \citep{1987MNRAS.227....1K} is implemented
through the $(1 + \beta_i \mu^2)^2$ factor, where $\beta_i =
f_{\Omega}/\avg{b_i}$ with $f_{\Omega} = \text{d}\ln{D}/\text{d}\ln{a}$ is the
logarithmic derivative of the growth factor, well approximated by $f_{\Omega}
\approx \Omega^{0.55}$ \citep{2005PhRvD..72d3529L}. We assume a Lorentzian form for the
finger-of-god suppression:
\beq\label{eq:fog}
D(x) = \frac{1}{1+x^2}\text{.}
\eeq
We approximate the pairwise velocity dispersion by $\sigma_p(z) =
\sigma_v(z)/\sqrt{2}$ with $\sigma_v^2(z)$ being the variance of the
line-of-sight component of the velocity field according to linear theory.
Equation~\eqref{eq:int_ps} neglects the shot noise component of the power
spectrum, which is an acceptable omission for the large scales considered in
this work. We use the \citet{1998ApJ...496..605E} model for the linear theory
matter power spectrum, $P_{\delta,\delta}(k, z)$. All basic cosmological
calculations are performed using the code \texttt{colossus} \citep{2018ApJS..239...35D}.

For the cross-spectrum between two lines $i$ and $j$, we assume that
\beq\label{eq:cross_spec}
P_{i,j}(k, \mu, z) = r_{i,j}(k, z) \sqrt{P_{i,i}(k, \mu, z) P_{j,j}(k, \mu, z)}\text{,}
\eeq
where $r_{i,j}$ is the cross-correlation coefficient between line $i$ and $j$.
We will only consider scales where $r_{i,j} = 1$.

\subsection{Multi-line Formalism}\label{ssec:multi_line}
Given a set of cross-spectra and their covariance matrix, we can determine the
optimal constraints that can be placed on both the average bias $\avg{b_i}$
and average intensity $\avg{I_i}$. For the set of parameters $\theta$
containing each $\avg{b_i}$ and $\avg{I_i}$, we can write down the Fisher
matrix as
\beq\label{eq:fisher}
F_{i,j} = 
\int \frac{\text{d}^3k}{(2\pi)^3} V_{\text{surv}}
% \int \frac{\text{d}k\,k^2}{2\pi^2} V_{\text{surv}} 
\sum_{l} \sum_{m}
\frac{\partial \hat{P}_{l}(\bm{k})}{\partial \theta_i}
\left(\Cov{\hat{P}_{l}(\bm{k})}{\hat{P}_{m}(\bm{k})}\right)^{-1}
\frac{\partial \hat{P}_{m}(\bm{k})}{\partial \theta_j}\text{,}
\eeq
where the sum over $l$ and $m$ each is over each \emph{pair} of lines. In
order to compute the covariance matrix, we use the result that \citep[e.g.][]{2010JCAP...11..016V},
\beq\label{eq:var_cov}
\begin{split}
\Var{\xps{i}{j}} &= \xps{i}{j}^2 + \pstot{i}\pstot{j} \\
\Cov{\xps{i}{j}}{\xps{i}{k}} &= \pstot{i}\xps{j}{k} +
\xps{i}{j}\xps{i}{k}\text{,}
\end{split}
\eeq
where $\pstot{i} = \ps{i} + N_i$. Note that $N_i$ contains contributions from
instrument noise, residual foreground contamination, and
interloping/extraloping lines.

While in principle one can extract $\bi$ and $\Ii$ independently due to the
$\mu$-dependence of the Kaiser effect, there are a number of observational
challenges \citep{2019arXiv190500209C}. We will therefore only consider the constraints
that can be placed on the product of the two, i.e., $B_i \equiv \bi \Ii$. The
previous Fisher matrix can be converted to the equivalent one for constraints
on $B_i$, $F'$, by the following relation \citep[e.g.][]{2009arXiv0906.4123C}:
\beq\label{eq:fisher_transform}
F' = M^{\text{T}} F M\text{,}
\eeq
where
\beq\label{eq:trans_matrix}
M_{ij} = \frac{\partial \theta_i}{\partial B_j}\text{.}
\eeq

>>>>

The intrinsic emission from a line $i$ is given by
\beq\label{eq:int_ps}
P_{i,i}(k, \mu, z) = \avg{I_i}^2 \avg{b_i}^2 (1 + \beta_i \mu^2)^2 
              D\left[\mu k \sigma_p(z)\right] P_{\delta, \delta}(k, z)\text{,}
              % + P_{i, \text{shot}}\text{,}
\eeq
where $\mu=k_{\parallel}/k$ is the cosine of the angle between the wavevector
$\bm{k}$ and the line of sight direction, $\avg{I_i}$ is the average specific
intensity of line $i$, and $\avg{b_i}$ is the average luminosity-weighted bias
of the emitting galaxies. The Kaiser effect \citep{1987MNRAS.227....1K} is implemented
through the $(1 + \beta_i \mu^2)^2$ factor, where $\beta_i =
f_{\Omega}/\avg{b_i}$ with $f_{\Omega} = \text{d}\ln{D}/\text{d}\ln{a}$ is the
logarithmic derivative of the growth factor, well approximated by $f_{\Omega}
\approx \Omega^{0.55}$ \citep{2005PhRvD..72d3529L}. We assume a Lorentzian form for the
finger-of-god suppression:
\beq\label{eq:fog}
D(x) = \frac{1}{1+x^2}\text{.}
\eeq
We approximate the pairwise velocity dispersion by $\sigma_p(z) =
\sigma_v(z)/\sqrt{2}$ with $\sigma_v^2(z)$ being the variance of the
line-of-sight component of the velocity field according to linear theory.
Equation~\eqref{eq:int_ps} neglects the shot noise component of the power
spectrum, which is an acceptable omission for the large scales considered in
this work. Our neglect to add the shot noise component is out of convenience considering the scales we will encounter, but this is not to imply that the shot noise component is uninteresting in its own right ({\bf cite some?}) even in the cross-spectra ({\bf cite Breysse 2019}). We use the \citet{1998ApJ...496..605E} model for the linear theory
matter power spectrum, $P_{\delta,\delta}(k, z)$. All basic cosmological
calculations are performed using the code \texttt{colossus} \citep{2018ApJS..239...35D}.

Throughout this work we will only consider the angle-averaged power spectrum, that is:
\beq\label{eq:int_ps}
\begin{split}
P_{i,i}(k, z) &= \avg{I_i}^2 \avg{b_i}^2 C_i(k) P_{\delta, \delta}(k, z)\\
                   &\equiv B_i^2 C_i(k) P_{\delta, \delta}(k, z)\text{,}
              % + P_{i, \text{shot}}\text{,}
\end{split}
\eeq
where we have defined $C_i(k)$ to be the amplification of the power spectrum
coming from the average over the angle-dependent terms in
Equation~\eqref{eq:int_ps}, and $B_i$ to be the product of $\avg{b_i}$ and
$\avg{I_i}$. Henceforth all power spectra will be assumed to be angle
averaged.

For the cross-spectrum between two lines $i$ and $j$, we assume that
\beq\label{eq:cross_spec}
P_{i,j}(k, z) = r_{i,j}(k, z) \sqrt{P_{i,i}(k, z) P_{j,j}(k, z)}\text{,}
\eeq
where $r_{i,j}$ is the cross-correlation coefficient between line $i$ and $j$.
We will only consider scales where $r_{i,j} = 1$.

\subsection{Multi-line Formalism}\label{ssec:multi_line}
Given a set of cross-spectra and their covariance matrix, we can, in
principle, determine the optimal constraints that can be placed on both the
average bias $\avg{b_i}$ and average intensity $\avg{I_i}$. However, such a
procedure requires measuring the angular dependence of the power spectrum
({\bf cite}), which may be fairly challenging \citep{2019arXiv190500209C}.
Therefore, after fixing the value of $b_i$ for each line $i$, we only consider
the constraint that can be placed on each $B_i$. This is slightly
inconsistent, since in principle we should first determine the constraint on
$\bi$ and $\Ii$ and then convert this into a constraint on $B_i$. But our
approximation is reasonable because {\bf of some reason}.

We can write down the Fisher matrix as
\beq\label{eq:fisher}
F_{i,j} = 
\int \frac{\text{d}^3k}{(2\pi)^3} V_{\text{surv}}
% \int \frac{\text{d}k\,k^2}{2\pi^2} V_{\text{surv}} 
\sum_{l} \sum_{m}
\frac{\partial \hat{P}_{l}(\bm{k})}{\partial B_i}
\left(\Cov{\hat{P}_{l}(\bm{k})}{\hat{P}_{m}(\bm{k})}\right)^{-1}
\frac{\partial \hat{P}_{m}(\bm{k})}{\partial B_j}\text{,}
\eeq
where the sum over $l$ and $m$ each is over each \emph{pair} of lines. Since
we will be considering only angle-averaged power spectra, we can use the
angle-averaged form of the Fisher matrix:
\beq\label{eq:fisher}
F_{i,j} = 
\int \frac{\text{d}k\,k^2}{2\pi^2} V_{\text{surv}}
% \int \frac{\text{d}k\,k^2}{2\pi^2} V_{\text{surv}} 
\sum_{l} \sum_{m}
\frac{\partial \hat{P}_{l}(k)}{\partial B_i}
\left(\Cov{\hat{P}_{l}(k)}{\hat{P}_{m}(k)}\right)^{-1}
\frac{\partial \hat{P}_{m}(k)}{\partial B_j}\text{.}
\eeq

In
order to compute the covariance matrix, we use the result that \citep[e.g.][]{2010JCAP...11..016V},
\beq\label{eq:var_cov}
\begin{split}
\Var{\xps{i}{j}} &= \xps{i}{j}^2 + \pstot{i}\pstot{j} \\
\Cov{\xps{i}{j}}{\xps{i}{k}} &= \pstot{i}\xps{j}{k} +
\xps{i}{j}\xps{i}{k}\text{,}
\end{split}
\eeq
where $\pstot{i} = \ps{i} + N_i$. Note that $N_i$ contains contributions from
instrument noise, residual foreground contamination, and
interloping/extraloping lines.
